<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{AUCM usage}
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE)
data.table::setDTthreads(1L)
```

# Maximizing AUC with the AUCM loss

AUCM is a loss function ported from LibAUC library.
It is a differentiable surrogate for AUC Loss specialized in imbalanced classification.

It adopts the min-max idea:
"a" should be the mean of the score of positive samples,
and "b" should be the mean of the score of negative samples.
The square error of "a" and "b" and their corresponding mean are minimized.

"alpha" is a dual variable which should be maximized.
"margin" is the desired gap of the positives and negatives.
"alpha" is a force that push the "a" and "b" apart to reach the value of the "margin".

Why?
First, mlr3torch's optimizer can't reach to the parameters in loss functions.
Second, "alpha", the dual variable, needs gradient ascent, which should be managed independently.

## Create an imbalanced version of spam data set

```{r}
spam_task <- mlr3::tsk("spam")
spam_task$filter(1504:4293) # for imbalance.  310    2480 
Class_vec <- spam_task$data(spam_task$row_ids, "type")$type
(count_tab <- table(Class_vec))
print(count_tab/sum(count_tab))
library(data.table)
orig_X_dt <- spam_task$data(spam_task$row_ids, spam_task$col_roles$feature)
sc_X_dt <- data.table(type=Class_vec, scale(orig_X_dt))
spam_sc_task <- mlr3::TaskClassif$new("spam_sc", sc_X_dt, target="type")
spam_sc_task$col_roles$stratum <- "type"
spam_list <- list(task=spam_sc_task)
```

## Run the Benchmark
```{r}
torch_available <- requireNamespace("mlr3torch") && torch::torch_is_installed()
if(torch_available){
  L <- mlr3torch::LearnerTorchMLP$new(task_type="classif")
  spam_dataset <- L$dataset(spam_sc_task)
  measure_list <- list(
    mlr3torchAUM::MeasureClassifInvAUC$new())
  batch_sampler_list <- list(
    random_full=mlr3torchAUM::batch_sampler_random(10000),
    random_stoch=mlr3torchAUM::batch_sampler_random(9))
  result_for_sampler <- function(batching){
    L <- mlr3torch::LearnerTorchMLP$new(task_type="classif")
    L$loss <- mlr3torchAUM::nn_AUCM_loss
    opt <- mlr3torch::as_torch_optimizer(mlr3torchAUM::optim_pesg)
    opt$param_set$set_values(
      lr = 0.01, clamp_value = 1, weight_decay = 1e-4,
      epoch_decay = 1e-3, momentum = 0.9, decay_factor = 2
    )
    L$optimizer <- opt
    L$predict_type <- "prob"
    L$callbacks <- list(mlr3torch::t_clbk("history"), 
    mlr3torchAUM::make_pesg_callback_full(
      lr = 0.01, clamp_value = 1, weight_decay = 1e-4,
      epoch_decay = 1e-3, momentum = 0.9, decay_factor = 2
    ))
    L$param_set$set_values(
      epochs=20, batch_size=32, seed=1,
      measures_valid=measure_list,
      measures_train=measure_list,
      batch_sampler=batch_sampler_list[[batching]])
    set.seed(1)
    mlr3::set_validate(L, 0.5)
    L$train(spam_sc_task)
    print(L$model$loss_fn)
    list(
      loss=sapply(L$model$loss_fn, torch::as_array),
      history=data.table(batching, L$model$callbacks$history),
      weights=torch::as_array(L$model$network$parameters[["0.weight"]]$flatten()))
  }
  result_list <- lapply(names(batch_sampler_list), result_for_sampler)
  names(result_list) <- names(batch_sampler_list)
  print(do.call(data.table, lapply(result_list, "[[", "weights")), nrow=10)
}
```