<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Compositional AUC usage}
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE)
data.table::setDTthreads(1L)
```

# Maximizing AUC with the Compositional AUC loss

`nn_CompositionalAUC_loss` is ported from LibAUC's `CompositionalAUCLoss`.
It builds on `nn_AUCM_loss`: instead of optimizing the AUC surrogate directly,
it optimizes

```
L_AUC(w - lr0 * grad L_CE(w))
```

Read that from the inside out.
The inner part takes one cross-entropy step from the current weights `w`.
The outer part evaluates the AUC-margin loss **at that moved position**
of the weights, not at `w`.
Composing them lets the second one work on top of a model the first one already
improved, one batch at a time.

In practice the two levels take turns.
With `k = 1` the schedule is CE, AUCM, CE, AUCM, ...;
with `k = 2` it is CE, CE, AUCM, AUCM, ....
The loss module counts its own forward calls and switches accordingly.

## Three pieces that must be used together

The loss carries three trainable parameters that live outside the network:
`a` (mean of the positive samplers), `b` (mean of the negative samplers) and
`alpha` (a dual variable that pushes `a` and `b` apart until they are `margin`
away from each other).
`mlr3torch`'s optimizer only sees the network parameters, so `a`, `b` and
`alpha` would never be trained by it.
That is why three things are needed at once:

| piece | trains |
|---|---|
| `nn_CompositionalAUC_loss` | holds `a`, `b`, `alpha` and the pass counter |
| `optim_pdsca` | the network weights, differently on each pass |
| `make_pdsca_callback` | `a`, `b`, `alpha`, and the per-epoch bookkeeping |

## Feed logits

Both passes need scores in `[0, 1]`.
By default `add_sigmoid = TRUE`, so the loss apply a sigmoid to
whatever the network emits before doing anything else,
and is suited for a normal `LearnerTorchMLP` whose head emits logits.

If your network already ends in a sigmoid, pass `add_sigmoid = FALSE`.

## Two learning rates, and a momentum convention that is backwards

`lr0` belongs to the inner cross-entropy step and `lr` to the outer AUC step;
they decay independently by `decay_factor0` and `decay_factor`.

The two momentum arguments are **not** PyTorch momentum:

```
buf <- (1 - beta) * buf + beta * new
```

`beta` is the weight on the *new* value, so a larger `beta` means *less*
smoothing, which is the opposite of what `optim_sgd(momentum=)` does.
`weight_momentum` averages the *weights* on cross-entropy batches,
while `momentum` averages the *gradients* on AUCM batches.

## An imbalanced version of the spam data set

```{r}
spam_task <- mlr3::tsk("spam")
spam_task$filter(1504:4293) # for imbalance:  310 vs 2480
Class_vec <- spam_task$data(spam_task$row_ids, "type")$type
(count_tab <- table(Class_vec))
print(count_tab / sum(count_tab))
library(data.table)
orig_X_dt <- spam_task$data(spam_task$row_ids, spam_task$col_roles$feature)
sc_X_dt <- data.table(type = Class_vec, scale(orig_X_dt))
spam_sc_task <- mlr3::TaskClassif$new("spam_sc", sc_X_dt, target = "type")
spam_sc_task$col_roles$stratum <- "type"
```

## Training

```{r}
torch_available <- requireNamespace("mlr3torch") && torch::torch_is_installed()
if(torch_available){
  hyper_params <- list(
    lr = 0.1, clamp_value = 1, weight_decay = 1e-4,
    epoch_decay = 1e-3, momentum = 0.9, decay_factor = 2)
  L <- mlr3torch::LearnerTorchMLP$new(task_type = "classif")
  L$loss <- mlr3torchAUM::nn_CompositionalAUC_loss
  opt <- mlr3torch::as_torch_optimizer(mlr3torchAUM::optim_pdsca)
  opt$param_set$set_values(
    lr0 = 0.05, lr = hyper_params$lr, clamp_value = hyper_params$clamp_value,
    weight_decay = hyper_params$weight_decay, epoch_decay = hyper_params$epoch_decay,
    weight_momentum = 0.9, momentum = hyper_params$momentum,
    decay_factor0 = 2, decay_factor = hyper_params$decay_factor)
  L$optimizer <- opt
  # the same hyperparameters again, this time for a, b and alpha
  L$callbacks <- do.call(mlr3torchAUM::make_pdsca_callback, hyper_params)
  L$predict_type <- "prob"
  L$param_set$set_values(
    epochs = 10, batch_size = 32, neurons = 8, shuffle = TRUE, seed = 1)
  set.seed(1)
  L$train(spam_sc_task)
  loss_fn <- L$model$loss_fn
  cat("a =", as.numeric(loss_fn$a), "\n")
  cat("b =", as.numeric(loss_fn$b), "\n")
  cat("alpha =", as.numeric(loss_fn$alpha), "\n")
  print(L$predict(spam_sc_task)$score(mlr3::msr("classif.auc")))
}
```

`a > b` is the sign that the loss parameters have separated the two classes the
way they should: positive scores concentrate above negative ones.
`alpha` is clamped at `0` from below, so a positive value means the margin has
not been reached yet and the dual variable is still pushing.

